Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/proof-reconstruction.tex
[ paper ] conjunct reviewed.
[ paper ] clausify added.
[ paper ] simplify and its example added.
[ paper ] final comment.

In the remainder part of this section, a formal description using
type theory is presented to build definitions and theorems of
functions necessary to emulate \Metis' inference rules.
As a convention, we write the type judgments as ``$m : M$'' to denote
that $m$ has $\abbre{M}$ type.
Types are written using small caps.
We use \Prop type for propositions and the dependent type, $Γ ⊢ φ$ for
theorems where $Γ$ has $\abbre{List}~\Prop$ type, and $φ$ has type $\Prop$.
$\abbre{List}$ type is the usual inductive type for lists.
For syntactical equivalence, we write the equality as $φ ≡ ψ$
for $φ : \Prop$ and $ψ : \Prop$.

@@

\begin{table}[!ht]
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &Thm.~\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &Thm.~\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &Thm.~\ref{thm:conjunct}
    \\
    \texttt{negate}
      &Negate the formula
      &Thm.~\ref{thm:negate}
    \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &Thm.~\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &Thm.~\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &Thm.~\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\label{tab:agda-metis-table}
\end{table}

@@

\subfile{sections/proof-reconstruction/strip.tex}
% \subfile{sections/proof-reconstruction/negate.tex}
% \subfile{sections/proof-reconstruction/resolve.tex}
% \subfile{sections/proof-reconstruction/canonicalize.tex}
% \subfile{sections/proof-reconstruction/conjunct.tex}
% \subfile{sections/proof-reconstruction/clausify.tex}
% \subfile{sections/proof-reconstruction/simplify.tex}

@@

>>>>

In the remainder part of this section, a formal description using
type theory is presented to build definitions and theorems of
functions necessary to emulate \Metis' inference rules.
As a convention, we write the type judgments as ``$m : M$'' to denote
that $m$ has $\abbre{M}$ type.
Types are written using small caps.
We use \Prop type for propositions and the dependent type, $Γ ⊢ φ$ for
theorems where $Γ$ has $\abbre{List}~\Prop$ type, and $φ$ has type $\Prop$.
$\abbre{List}$ type is the usual inductive type for lists.
For syntactical equivalence, we write the equality as $φ ≡ ψ$
for $φ : \Prop$ and $ψ : \Prop$. For this equality, we assume the
reflexivity, symmetry and transitivity property.

@@

\begin{table}[!ht]
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline
    \texttt{canonicalize}
      &Normalization of the formula
      &Thm.~\ref{thm:canonicalize}
    \\
    \texttt{clausify}
      &Performs clausification
      &Thm.~\ref{thm:clausify}
    \\
    \hyperlink{atp-conjunct}{\texttt{conjunct}}
      &Takes a formula from a conjunction
      &Thm.~\ref{thm:conjunct}
    \\
    % \texttt{negate}
    %   &Negate the formula
    %   &Thm.~\ref{thm:negate}
    % \\
    \texttt{resolve}
      &A general form of the resolution theorem
      &Thm.~\ref{thm:resolve}
    \\
    \texttt{simplify}
      &Simplify definitions and theorems
      &Thm.~\ref{thm:simplify}
    \\
    \texttt{strip}
      &Strip a goal into subgoals
      &Thm.~\ref{thm:strip}
    \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\label{tab:agda-metis-table}
\end{table}

@@

\subfile{sections/proof-reconstruction/strip.tex}
\subfile{sections/proof-reconstruction/negate.tex}
\subfile{sections/proof-reconstruction/conjunct.tex}
% \subfile{sections/proof-reconstruction/resolve.tex}
% \subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}

@@

Now, we have finished the formalization of every inference rule in 
a \Metis proof, we are able to justify step-by-step any proof
for a problem in propositional logic. 
For instance, we tested successfully the translation by \Athena jointly with 
the \Agda formalizations of the rules mentioned above with more than eighty representative theorems in propositional logic. An interested reader can test the problems~\cite{Prieto-Cubides2017} in the \Athena tool repository~\cite{Athena}.
